\LARGE \bf Explicit Knowledge and the Deliberative Layer: Lessons Learned

### 2013-03-13 paper.tex
Added a couple of macros
Completed the 'Communication' section

\usepackage[utf8x]{inputenc}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\graphicspath{{figs/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

@@

@@

\cite{Lemaignan2011a}

@@

One protoypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child's \emph{spygame}: two players are facing
eachother with a set of random objects inbetween, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?}. Based on the knowledge it
has acquired, the robot is able to minimize the number of questions required to
find the object.

@@

>>>>

\usepackage[protrusion=true,expansion=true]{microtype}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\graphicspath{{figs/}}
\DeclareGraphicsExtensions{.pdf,.jpg,.png}

\usepackage{ifthen}

\usepackage[utf8x]{inputenc}

@@

\usepackage{listings}
\usepackage{alltt}
\usepackage{fancyvrb}

\newcommand{\concept}[1]{{\small \texttt{#1}}}

\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
%\newcommand{\stmt}[1]{{\footnotesize $\langle\stmttt#1\relax\rangle$}}
%\newcommand{\rawstmt}[1]{{\footnotesize \stmttt#1\relax}}
%\def\stmttt#1 #2 #3\relax{
%\texttt{\IfBeginWith{#1}{?}{\textbf{#1}}{#1} \IfBeginWith{#2}{?}{\textbf{#2}}{\emph{#2}} \ifthenelse {\equal{#3}{true} \OR \equal{#3}{false}}{\emph{#3}}{\IfBeginWith{#3}{?}{\textbf{#3}} {#3}}}}

%\newcommand{\setstmt}[1]{{\footnotesize [\setstmttt#1\relax]}}
%\def\setstmttt#1,#2\relax{\stmttt#1\relax, \stmttt#2\relax}

@@

One of the field of human-robot interaction where the introduction of the
semantic layer has been most beneficial is natural language processing.
\cite{Lemaignan2011a} details the techniques and the tool called
\texttt{dialogs} that we have developed for English language parsing and
grounding, for (minimalist) dialogue management and for verbalisation.

Natural language input (in experiments, we rely on an Android-based interface
running on a tablet, with Google-based speech recognition, for the input) is
parsed into a grammatical structure, and each sentence atoms are resolved with
the help of the ontology to ground classes (when a user says ``pick the can'',
to which instance of \emph{can} the user refers to?) and actions. Sentences are
sorted into questions, desires and statements, and processed accordingly.

The system supports quantification (``give me \{a|the|some|all|any|...\} can''),
thematic roles attached to actions, interactive disambiguation (the robot asks
questions when needed), anaphora resolution (``give \emph{it} to me'') based on
dialogue history. It also permits knowledge extension by learning new semantic
structures (for instance, a sentence like ``learn that cats are animals'' is
converted into \stmt{Cat subClassOf Animal}, even if \concept{Cat} is not known
beforehand) and tries to translate \emph{states} (``I'm tired'') into
\emph{experiences} (\stmt{HUMAN experiences state\_1, state\_1 hasFeature
tired}).

@@

One prototypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child's \emph{spygame}: two players are facing
each other with a set of random objects in-between, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?}. Based on the knowledge it
has acquired, the robot is able to minimize the number of questions required to
find the object.

@@

Because the main communication channel between component is semantic data,
different communication modalities (explicit like verbal, dectic or based on
gaze, or implicit like postures) are accessed in an homogeneous way.

The grounding process copes with and integrates the different physical
modalities at two disctint levels.

First, the grounding module explicitely check for the presence and value of
particular facts: for instance, when a set of instances match a category (the
human says ``give me the bottle'' and the robot knows about three bottles), the
module may decide (it actually depends on the quantifier preceding the class:
\emph{the}, \emph{a}, \emph{any}...) to discard some of them based on their
\emph{visibility} by the speaker (implicit communication context built from the
human posture).

Another example, when the human says ``this'', the robot checks if the human is
currently both pointing and looking at some object. In that case, \emph{this}
is replaced by the object focused on.

Note that, while the system benefits from the complementary modalities, they
are not required. The dialogue system would work as well in a purely verbal
setting.

The second level of integration of multi-modality is implicit: by computing
symbolic properties from the geometry, richer descriptions and hence
discrimination possibilities are available: for instance, if reachability is
available, the robot may asks ``do you mean the bottle that is accessible to
me?'' to discriminate between the three bottles. That way, procedures relying
on discrimination transparently benefit from added modalities.
